Mining Interaction Log Data in a Creative Arts Practice using Macro Gestures and Transition Matrices

### 2015-09-21 flux-entropy-paper/flux-entropy-paper.tex
a few more macro gesture fixes

In our corpus of iPad musical performances, rehearsals and studies, we
have used an automated method to classify logs of touch interaction,
at a regular interval, into nine continuous touch-screen gestures, as
described by Martin et al~\cite{Martin:2015jk}. Each performer in each
musical piece can then be thought of as producing a ``score'' of state
transitions and an overall transition matrix for the entire ensemble
can be derived by averaging the individual transition matrices.

Suppose that each participant's activity in a collaborative
interaction can at any given time be classified as belonging to one of
$m$ distinct \emph{states}. Over the course of the activity, each
participant $X$ generates a length $N$ sequence of these interaction
states $X_n$ ($n = 1, \ldots, N$) which can be thought of as Markov
process.

@@

We applied a Random Forest~\cite{Breiman:2001kx}
classifier\footnote{The Random Forest classifier used was from
  Python's Scikit-learn package\cite{scikit-learn}.} to 5-second
windows of touch-data in our corpus to provide chains of
gesture-states for each participant in each session following previous
work~\cite{Martin:2015jk}. These chains can be used to create
$9 \times 9$ (the classifier recognises 9~distinct gestures)
transition matrices for individual performers and the whole ensemble
in each session.

@@

While the interaction sessions recorded in our corpus were performed
by many different participants with varying experience levels, in
different contexts and on different instrumental setups, the common
interface and method of recording the data allows all the sessions to
be transformed into gesture sequences. These transition matrices have
been used as a visualisation technique to reveal qualitative
differences in structure in improvisational
music-making\cite{Swift:2014tya}. In this paper, however, we wish to
extract quantitative statistics from these TMs for comparing and
tracking different sessions over time.

>>>>

In our corpus of iPad musical performances, logs of low-level touch
data record each performers micro-gestures. However, it has been
established that musical performances on touch-screens can be seen in
terms of vocabularies of continuous macro gestures that unfold over
several seconds of touch interaction~\cite{Martin:2014cr}. Improvised
performances can be then thought of as a sequence of such
states~\cite{Pressing:1988uo} for each performer.

If the performers' touch interactions can be classified as belonging
to one of $m$ distinct states, then over the course of a session each
participant $X$ generates a length $N$ sequence of these interaction
states $X_n$ ($n = 1, \ldots, N$) which can be thought of as Markov
process.

@@

To generate state sequences for our corpus we have used an automated
method to classify logs of touch interaction, at a regular interval,
into nine continuous touch-screen gestures, as described by Martin et
al~\cite{Martin:2015jk}. A Random Forest~\cite{Breiman:2001kx}
classifier\footnote{The Random Forest classifier used was from
  Python's Scikit-learn package\cite{scikit-learn}.} was applied to
5-second windows of touch-data in our corpus to provide chains of
gesture-states for each participant in each session. These chains can
be used to create $9 \times 9$ transition matrices for individual
performers and the whole ensemble in each session.

@@

While the interaction sessions recorded in our corpus were performed
by many different participants with varying experience levels, in
different contexts and on different instrumental setups, the common
interface and method of recording the data allows all the sessions to
be transformed into gesture sequences, and transition matrices.
Transition matrices have been used as a visualisation technique to
reveal qualitative differences in structure in improvisational
music-making\cite{Swift:2014tya}. In this paper, however, we wish to
extract quantitative measures from these TMs for comparing and
tracking different sessions over time.
